% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{section title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{subsection title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\geometry{margin=1in}
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering


\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
\usepackage{caption}
% Make caption package work with longtable
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{geometry}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Curriculum Vitae},
  pdfauthor={Steven Mesquiti, M.A.},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}


\title{Curriculum Vitae}
\author{Steven Mesquiti, M.A.}
\date{}

\begin{document}
\frame{\titlepage}


\begin{frame}{Steven Mesquiti, M.A.}
\phantomsection\label{steven-mesquiti-m.a.}
Department of Psychology\\
Princeton University\\
sm9518@princeton.edu
\end{frame}

\begin{frame}
\begin{block}{Education}
\phantomsection\label{education}
\textbf{Princeton University} (2024-Present)\\
Ph.D.~in Psychology\\
Advisor: Dr.~Erik Nook

\textbf{Texas State University} (2019-2022)\\
M.A.~in Psychological Research\\
Advisor: Dr.~Randall Osborne

\textbf{Southwestern University} (2015-2019)\\
B.A. in Psychology (Minor: Spanish)
\end{block}
\end{frame}

\begin{frame}
\begin{block}{Publications}
\phantomsection\label{publications}
\begin{itemize}
\tightlist
\item
  Denotes shared first authorships
\end{itemize}

\textbf{Mesquiti, S., Cosme, D., Nook, E.C., Falk E.B., Burns S.} (Under
Review). \emph{Predicting Psychological and Subjective Well-being
through Language-based Assessments of Well-being.}

\textbf{Mesquiti, S., Seraj, S., Weyland, A. H., Ashokkumar, A., Boyd,
R. L., Mihalcea, R., \& Pennebaker, J. W.} (2025). \emph{Analysis of
social media language reveals the psychological interaction of three
successive upheavals.} \emph{Scientific Reports, 15}(1), 5740.
\href{https://doi.org/10.1038/s41598-025-89165-z}{DOI}.

\textbf{Kang, Y., Mesquiti, S., Baik, E.S., \& Falk, E. B.} (2024).
\emph{Empathy and Helping: The Role of Affect in Response to Others'
Suffering.} \emph{Scientific Reports, 15}, 3256
\href{https://doi.org/10.1038/s41598-025-87221-2}{DOI}.

\textbf{Mesquiti, S., Seraj, S.} (2023). \emph{The Psychological Impacts
of the COVID-19 Pandemic on Business Leadership.} \emph{PLOS One.}
\href{https://doi.org/10.1371/journal.pone.0290621}{DOI}
\end{block}
\end{frame}

\begin{frame}
\begin{block}{Works in Progress}
\phantomsection\label{works-in-progress}
\begin{itemize}
\tightlist
\item
  \textbf{Kang, Y., Mesquiti, S. \& Falk, E. B.} (in prep). \emph{The
  Effect of Compassion Training on Emotional Support Giving.}
\item
  \textbf{Mesquiti, S. \& Nook, E.C.} (in prep.) \emph{The Implication
  of Artificial Intelligence on Mental Health.}
\item
  \textbf{Mesquiti, S., Stade E. C., Hull, T.D., Nook, E. C.} (in prep.)
  \emph{Tracking Shifts in the Latent Meaning of ``I'' and Its
  Connection to Mental Health Outcomes in a Large Therapy Dataset.}
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
\begin{block}{Funding}
\phantomsection\label{funding}
\begin{itemize}
\tightlist
\item
  \textbf{2025}: New Jersey Health Foundation: \emph{Investigating
  Racial Equity of Psycholinguistic Measures of Mental Health} (PI: Erik
  C. Nook; \$34,867; \# PC 42-25)
\item
  \textbf{2022}: Northwestern University Kellogg School of Management
  Behavioral Research Fellowship (Finalist {[}Declined{]}; \$35,000)
\item
  \textbf{2020}: Texas State University Thesis Research Support
  Fellowship (\$1,070)
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
\begin{block}{Honors \& Awards}
\phantomsection\label{honors-awards}
\begin{itemize}
\tightlist
\item
  \textbf{2024-25}: Princeton University William G. Bowen Merit
  Fellowship
\item
  \textbf{2020-21}: Texas State University Graduate College of Liberal
  Arts Scholarship
\item
  \textbf{2021}: Masters in Psychological Research Graduate Competitive
  Scholarship
\item
  \textbf{2019}: Texas State University Graduate College of Education
  Scholarship
\item
  \textbf{2015}: Southwestern University Mood Scholar
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
\begin{block}{Invited Talks}
\phantomsection\label{invited-talks}
\begin{itemize}
\tightlist
\item
  \textbf{2025}: \emph{Tracking Shifts in the Latent Meaning of ``I''
  and Its Connection to Mental Health Outcomes in a Large Therapy
  Dataset.} Social Psychology Seminar, Princeton University, Princeton,
  NJ.
\item
  \textbf{2022}: \emph{Within-and between-person effects of autonomous
  motivation on goal pursuit.} Duckworth Lab, Princeton University.
\item
  \textbf{2021}: \emph{The whole truth and nothing but the truth: an
  analysis of the variability of prosocial lying in the Netherlands and
  the United States.} Texas State University Psychology Department Brown
  Bag.
\item
  \textbf{2021}: \emph{The truth about prosocial lying: Cross-cultural
  differences in prosocial lying.} Southwestern Psychological
  Association (Virtual).
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
\begin{block}{Conference Presentations}
\phantomsection\label{conference-presentations}
\emph{Denotes Undergraduate Mentored}

\begin{itemize}
\tightlist
\item
  \textbf{Mesquiti, S. C., Nook, E.} (2024, March). \emph{Investigating
  Racial Sensitivity in Language-Based Assessments of Mental Health.}
  Poster presented at the annual meeting for the Society for Affective
  Science.
\item
  \textbf{Mesquiti, S. C., Cosme, D., Nook, E., Falk E., Burns S.}
  (2025, February). \emph{Predicting Well-being with Language-based
  Assessments.} Poster presented at the annual meeting for the Society
  for Personality and Social Psychology.
\item
  \textbf{Mesquiti, S. C., Mobasser, A., Falk, E., Pfeifer, J. H., \&
  Cosme, D} (2023, February). \emph{Within-and between-person effects of
  autonomous motivation on goal pursuit.} Poster submitted to the annual
  meeting for the Society for Personality and Social Psychology.
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
\begin{block}{References}
\phantomsection\label{references}
\begin{itemize}
\item
  \textbf{Emily Falk}\\
  Professor of Communication, Psychology, Marketing, and OID
  (Operations, Informatics, and Decisions), Vice Dean of the Annenberg
  School for Communication, University of Pennsylvania\\
  efalk@falklab.org
\item
  \textbf{James Pennebaker}\\
  Professor Emeritus of Psychology, The University of Texas at Austin\\
  pennebaker@utexas.edu
\item
  \textbf{Erik Nook}\\
  Assistant Professor of Psychology, Princeton University\\
  enook@princeton.edu
\end{itemize}
\end{block}
\end{frame}




\end{document}
